---
title: "Pink Salmon Vessel Calibration Coefficient Discussion"
author: "Sara Miller, Rich Brenner, Jim Murphy"
date: "November 4, 2020--draft"
#output:
#  word_document: default
#  html_document: default
#  pdf_document: default


output:
  bookdown::pdf_document2:
    fig_caption: yes
    toc: yes
header-includes:
 \usepackage{float}   
---
```{r setup, echo = FALSE, warning = FALSE, message = FALSE, error = FALSE}
library(here)
library(fs)
library(tidyverse)
library(knitr)
library(ggplot2)
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message = FALSE, error=FALSE)
year.forecast <- "2021_forecast" 
index <- "index4" #change based on index
data.directory <- file.path(year.forecast, index, 'data', '/')
results.directory <- file.path(year.forecast, index, 'results', '/')
```

```{r data}

fs::dir_ls(here::here(results.directory), regexp = "\\.csv$") %>% 
  set_names(nm = (basename(.) %>% 
                    tools::file_path_sans_ext())) %>%
  map(read_csv) -> tbls_seak

```
# Objective
To determine the vessel calibration coefficient to calculate the corresponding index of juvenile abundance for the 2021 SEAK pink salmon forecast models.

# Background
Excerpted from Wertheimer et al. 2010:  

"From 1997 to 2007, SECM used the NOAA ship *John N. Cobb* to accrue an 11 year time series of catches with a Nordic 264 rope trawl fished at the surface... (Orsi et al. 2000, 2008)... In 2007, in anticipation of the decommissioning of the *John N. Cobb*, the *Medeia* and the *John N. Cobb* fished synoptically for 28 pairs of trawl hauls to develop calibration factors in the event of differential catch rates between the two vessels (Wertheimer et al. 2008).  In 2008, the *Medeia* fished synoptically with the chartered research vessel *Steller* to determine relative fishing efficiency so that *Steller* catches could then be compared and calibrated to the SECM data series from the *John N. Cobb* (Wertheimer et al. 2009).  In 2009, the commercial trawler *Chellissa* was chartered to fish the SECM transects in the northern and southern regions of Southeast Alaska. The *Medeia* was again fished synoptically in the northern region transects to determine relative fishing efficiency (Table 1)."

Historically....  
For the 2021 SEAK pink salmon forecast, there was a discussion as to which vessel calibration coefficient to use going forward. Using the four potential vessel calibration types (pink_cal_mixspecies, pink_cal_mixpool, pink_cal_species, pink_cal_pool; Table 1), the corresponding index of juvenile abundance was slightly different (i.e., CPUE; standardized pink salmon catch based on a 20 minute trawl set by year; Table 2). To calculate the index of juvenile abundance by calibration type, the log-transformed pink salmon catch (standardized to an effort of a 20 minute trawl set), by haul, is multiplied by a vessel calibration coefficient (Table 1). Then, this value is averaged by month and year; whichever month had the highest average catches in a given year is then used as the juvenile abundance index for that particular year (Table 2). This index of juvenile abundance is then used as the variable 'CPUE' in the 2021 SEAK pink salmon forecast models.

```{r coefficients1}
tbls_seak$calibration_coeff %>% 
   knitr::kable(format = 'pandoc', caption = 'Vessel calibration coefficients (inverse of the fishing power coefficients in Table 3) used to convert vessel-specific catches to *Cobb* units.', row.names = F)
```

```{r coefficients2}
tbls_seak$peak_values %>% 
   knitr::kable(format = 'pandoc', caption = 'The datasets for the variable CPUE (index of juvenile pink salmon abundance based on log-transformed catches that are standardized to an effort of a 20 minute trawl set) using different vessel calibration coefficients.', row.names = F)
```

# Vessel Calibration Coefficients
The four types of vessel calibration coefficients are defined as:  

* pink_cal_mixspecies is a mixture of species-specific (*Chellisa:Medeia*) and pooled-species (*Medeia:Cobb*) coefficients for the *Chellissa* and the *N/W Explorer* (i.e., 1/(1.19 x 1.27) where 1.19 is the pooled-species fishing power coefficient from the *Medeia:Cobb* and 1.27 is the species-specific fishing power coefficient from the *Chellisa:Medeia*; see Table 3), and species-specific coefficients for the *Steller* and the *Medeia*. This ends up with a time series that is based on coefficients that vary with species, but are partially derived from a mixture of species-specific and pooled-species coefficients.  

* pink_cal_mixpool is a mixture of species-specific (*Chellissa:Medeia*) and pooled-species (*Medeia:Cobb*) coefficients for the *Chellissa* and the *N/W Explorer* (i.e., 1/(1.19 x 1.27)), and pooled-species coefficients for the *Steller* and the *Medeia*.

* pink_cal_species uses species-specific coefficients for the *Chellisa*, *Steller*, and *Medeia*,  and the *N/W Explorer* uses a species-specific coefficient from the *Chellisa*.

* pink_cal_pool uses pooled-species coefficients for the *Chellisa*, *Steller*, and *Medeia*, and the *N/W Explorer* uses a pooled-species coefficient from the *Chellisa*.

The term species-specific refers to a pink salmon specific vessel calibration coefficient, and the term pooled-species refers to a pink, chum, sockeye, and coho (i.e., catch of all species summed) vessel calibration coefficient.

The four vessel calibration coefficients (Table 1) are calculated as the inverse of the fishing power coefficients (Table 3). 

```{r coefficients3}
tbls_seak$fishing_power %>% 
   knitr::kable(format = 'pandoc', caption = 'Estimated fishing power coefficients for juvenile salmon catches by the different vessels used during the Southeast Alaska Coastal Monitoring survey (Wertheimer et al.  2008, 2009, and 2010). One of the primary trawl vessels, F/V *Northwest Explorer*, has not been calibrated and it is assumed to have the same fishing power as the *Chellissa*.', row.names = F)
```

Table 4 (below) shows how the vessel calibration coefficients (in Table 1) are calculated by taking the inverse of the fishing power coefficients in Table 3. 

```{r coefficients4}
tbls_seak$calibration_coeff_calc %>% 
   knitr::kable(format = 'pandoc', caption = 'Specific calculations for determining the vessel calibration coefficients  from the fishing power coefficients.', row.names = F)
```

# Analysis
Performance metrics (Akaike Information Criterion corrected for small sample sizes; AICc values; Burnham and Anderson 2004; mean and median absolute percentage error (MAPE, MEAPE); mean absolute scaled error (MASE) (Hyndman and Kohler 2006)) were used to evaluate forecast accuracy of four alternative vessel calibration coefficients (Table 1) using the same model. For all of these metrics, the smallest value is the preferred model. The model used for the comparison of the vessel calibration coefficients was:

$$E(y) = \alpha + \beta_1{X_1} + \beta_2{X_2},$$   

where ${X_1}$ is CPUE (i.e., juvenile pink salmon abundance index based on one of the four different vessel calibration coefficients; Table 2), and ${X_2}$ is the average temperature in Icy Strait in May, June, and July at the eight SECM stations in Icy Strait. CPUE data are log-transformed. Statistical analyses were performed with the R Project for Statistical computing version 3.6.3 (R Core Team 2020).


```{r coefficients5}
tbls_seak$model_compare %>% 
   knitr::kable(format = 'pandoc', caption = 'Comparison of the performance metrics for a model based on calculating a time-series of juvenile pink salmon abundance using different vessel calibration coefficients.', row.names = F)
```

# Conclusion
Based on performance metrics, the pooled-species vessel calibration coefficient was preferred (Table 5). Therefore, the index of juvenile abundance for the 2021 SEAK pink salmon forecast will be based on the pooled-species vessel calibration coefficient.

# References
Burnham, K. P., and D. R. Anderson. 2004. Multimodel inference: Understanding AIC and BIC in model selection. Sociological Methods and Research 33: 261-304.

Hyndman, R. J. and A. B. Koehler. 2006. Another look at measures of forecast accuracy. International Journal of Forecasting 22: 679-688.

Orsi, J. A., M. V. Sturdevant, J. M. Murphy, D. G. Mortensen, and B. L. Wing. 2000. Seasonal habitat use and early marine ecology of juvenile Pacific salmon in southeastern Alaska. NPAFC Bull. 2:111-122.

Orsi, J. A., E. A. Fergusson, M. V. Sturdevant, B. L. Wing, A. C. Wertheimer, and W. R. Heard. 2008. Annual survey of juvenile salmon and ecologically related species and environmental factors in the marine waters of southeastern Alaska, May–August 2007. NPAFC Doc. 1110, 82 pp.

R Core Team. 2020. R: A language and environment for statistical computing. R Foundation for Statistical Computing, Vienna, Austria. URL: http://www.r-project.org/index.html 

Wertheimer, A. C., J. A. Orsi, E. A. Fergusson, and M. V. Sturdevant. 2008. Paired comparisons of juvenile salmon catches between two research vessels fishing Nordic 264 surface trawls in southeastern Alaska, July 2007. NPAFC Doc. 1112.,17 p.

Wertheimer, A. C., J. A. Orsi, E. A. Fergusson, and M. V. Sturdevant. 2009. Calibration of Juvenile Salmon Catches using Paired Comparisons between Two Research Vessels Fishing Nordic 264 Surface Trawls in Southeastern Alaska, July 2008. NPAFC Doc. 1180. 18 pp.

Wertheimer, A. C., J. A. Orsi, E. A. Fergusson, and M. V. Sturdevant. 2010. Calibration of Juvenile Salmon Catches using Paired Comparisons between Two Research Vessels Fishing Nordic 264 Surface Trawls in Southeast Alaska, July 2009. NPAFC Doc. 1177. 19 pp. (Available at http://www.npafc.org).

```{r sess_info, echo=FALSE}
#sessionInfo()
```
